### 2015-06-19 paper.tex
Fixed some typos too.

\documentclass[fleqn,10pt,twoside]{gcb15submission}
\usepackage{url}\urlstyle{same}
\usepackage{booktabs}
\usepackage{colortbl, xcolor}

% rysunki
\usepackage{tikz}
\usepackage{ifthen}
\usepackage{xxcolor}
\usetikzlibrary{arrows}
\usetikzlibrary[topaths]
\usetikzlibrary{decorations.pathreplacing}

\title{signalHsmm - a novel semi-Markov model of eukaryotic signal peptides}

\author[1]{Micha\l{}  Burdukiewicz}
\author[2]{Piotr Sobczyk}
\author[1]{Pawe\l{} B\l{}a\.{z}ej}
\author[1]{Pawe\l{} Mackiewicz}
\affil[1]{University of Wroc\l{}aw, Department of Genomics, Poland}
\affil[2]{Wroc\l{}aw University of Technology, Department of Mathematics, Poland}

\keywords{signal petide prediction; n-gram; hidden semi-Markov models}

\begin{abstract}
The proper localization of proteins in a cell is essential to maintain their desired function. Information about the protein destination is included within the very protein in the form of short peptides called targeting signals. Ones of them are signal peptides, diverse N-terminal sequences, which are responsible for targeting of proteins to endomembrane system and their export outside the cell. Proteins equipped with signal peptides play crucial roles in metabolism, maintenance of tissue structure, immune response and regulation of other organismal functions.

A common model of classical signal peptides assumes that they start with a positively charged n-region, followed by a hydrophobic h-region and a c-region ended with a cleavage site recognized by a signal peptidase. However, studies of many protein sequences representing the wide range of diversified taxonomic organisms indicate a variability of signal peptides. Therefore, we designed a new, more universal probabilistic model for eukaryotic signal peptides, which includes knowledge about their organization, amino acid composition and variation.

The proposed model is based on hidden semi-Markov models (HSMMs) and use intrinsic knowledge about signal peptides. The big advantage of the algorithm is its extensibility. Using the n-grams (k-mers) we point how the general model can be attuned to yield not only better results, but also more information about signal peptides.

Our model was validated as a signal peptide predictor. It has showed the largest AUC=0.97 in comparison to other software and appeared very stable in the recovery of signal peptides after training even on very small data sets. Thanks to that, our model does not need to be permanently retrained with the continuous expansion of sequence databases. It should be emphasized that our model describes signal peptides from medically significant malaria parasites Plasmodium and their relatives (AUC = 0.92) more accurately than popular programs (0.84).

The web-server of signalHsmm is avaible under address \url{http://smorfland.uni.wroc.pl/signalhsmm}.

\end{abstract}

\begin{document}
\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}

\subsection*{Signal peptides}

Proteins of eukaryotes are encoded in nuclear genomes and are synthesized in ribosomes located in the cytosol or bounded by the endoplasmic reticulum. After the translation process, proteins have to be targeted to specific subcellular compartments or exported outside the cell to the extracellular environment. The proper localization of proteins is essential to perform their desired function. Information about protein destination is included within the very protein in the form of short peptides or stretches of amino acid residues called targeting or sorting signals. Ones of them are signal peptides, which are located at the N-terminus of proteins.

Signal peptides are responsible for targeting of proteins via the Sec61 translocation channel~\citep{2007rapoportprotein} to endomembrane system, which includes endoplasmic reticulum, Golgi apparatus and endosomes. Such proteins can stay inside of these compartments, or can be inserted into cellular membranes or exported outside the cell. Proteins equipped with signal peptides constitute a substantial fraction of the whole proteome. They play crucial roles in metabolism ($\beta$ galactosidase, pepsins)~\citep{1991hofmannmutations}, maintenance of tissue structure (collagen)~\citep{2001chanaberrant}, immune response (interferons, interleukins)~\citep{2005zhangalteration} and regulation of other organismal functions (prolactin, glucagon)~\citep{2010huangrole}. Moreover, passing proteins through the endomembrane system is important for their correct folding and posttranslational modification such as glycosylation and phosphorylation.

Despite the low sequence homology between signal peptides~\citep{1999ladungaphysean}, some general architecture were proposed~\citep{1994izardsignal, 2013vossmechanism} - Fig. \ref{fig:sparch}. It is assumed that signal peptides start with a positively charged sequence of amino acid residues, called the n-region with the length of about 5-8 residues. They probably enforce a proper topology on the polypeptide during translocation through membrane based on the positive-inside rule~\citep{1988vonheijnetopogenic}. The first region is followed by a stretch of hydrophobic amino acids (h-region) with the length of about 8-12 residues. It constitutes a core region of signal peptide and has a tendency to form $\alpha$-helix. The third part of a signal peptide, usually 6 residues long, is a polar, but uncharged c-region ended with a cleavage site recognized by the signal peptidase. The amino acid composition and the length of these regions vary between signal peptides, which influences the efficiency of protein secretion~\citep{2006hegdethe}.

\begin{figure}[ht]\centering
\includegraphics[width=0.55\textwidth]{figures/SP.png}
\caption{The organization of signal peptide. The figure is not drawn to scale.}
\label{fig:sparch}
\end{figure}

During or after translocation of the protein into the lumen of endoplasmic reticulum, the typical signal peptide is cleaved by a signal peptidase~\citep{2002paetzelsignal} and next degraded by specific proteases, whereas the rest (mature) part of protein stays in the lumen or is passed to other compartments. The cleavage site is characterized by a very variable amino acid composition. It typically contains small and neutral residues at -3 and -1 positions~\citep{1994palzkillselection}. The site is, however, absent from some membrane proteins in which the first transmembrane-domain acts both as signal peptides and signal anchor~\citep{1988szczesnaskorupapositive}.

Some data indicate that signal peptides may be universal. It was found, for example, that even bacterial signal peptides targeted correctly transgenic proteins to the plant~\citep{2009moellera} or mammalian secretory system~\citep{2014naganoestablishment}. On the other hand, signal peptides show great variation and the description presented above (Fig. \ref{fig:sparch}) refers to the most ‘typical’ signal peptides. There are also exceptionally long signal peptides, which fulfill more sophisticated roles~\citep{2009hissarchitecture}. For example, the fragment of signal peptide from preprolactin probably takes part in the regulation of prolactin secretion. Other examples are signal peptides of MHC class I, which inhibit activity of NK cells. Interesting functions have signal peptides of viral origin, which are involved in the immune evasion or viral life cycle~\citep{2000kappposttargeting}. Such diverse functions are restricted not only to the long signal peptides. The peptides from midkine, a protein contributing to the tumor progression, contains epitopes recognized by CD4+ T cells~\citep{2013kerzerhothe}. It indicates that the signal peptide may participate in a tumor immunity. The functional significance of these targeting signals makes that the prediction of signal peptide-containing proteins is also an important step in the drug development~\citep{2005zhangalteration, 2012netoadeimproving, 2010moellerwetmilling}.

\subsection*{Signal peptide predicting software}

Although many experimental methods determining the subcellular localization of proteins were devised, they are time consuming and laborious. Therefore, the development of new approaches in the field of computational biology and bioinformatics is desirable. They are not only a good support, complement or alternative for the experimental methods but also enable to understand rules encoding information about protein targeting, which is contained in the predicted targeting signals. However, many of the present predictors disregard full biological information carrying by signal peptides.

Having functional importance and specific features, signal peptides became the subject of many programs to their prediction based on different methods. The state of the art software for predicting the presence of signal peptides often incorporates ‘black-box’ models, as neural networks~\citep{2011petersensignalp}, support vector machines~\citep{2014zhangprediction}, Bayesian networks~\citep{2012zhengsignalbnf} or k-nearest neighbours~\citep{2007shensignall}. Such models neither utilize biological information about signal peptides nor work properly on atypical signal peptides.

The other type of software incorporates into the prediction process theoretical knowledge about the structure of signal peptides. Although these programs do not share innate flaws of ‘black-box’ models, they also demand an improvement. Some of them are based on position matrices or their more abstract variants~\citep{2014zhangprediction, 2004hillerpredisi}. Others (Phobius, Philius and not supported SignalP 3.0) use hidden Markov models (HMMs)~\citep{2004klla, 2008reynoldstransmembrane, 2004bendtsenimproved}, which reflect regions of signal peptides without realizing limitations of this probabilistic framework. However, the HMMs imply geometric distribution for duration of region length. We replicated the rules for extracting regions' boundaries from the first work utilizing HMMs in prediction of signal peptides~\citep{1998nielsenprediction} and found that the mentioned assumption is not in concordance with the reality because every region revealed the length distribution other than geometric (Fig. \ref{fig:reglen}). Moreover, the commonly used rigid region scheme (Fig. \ref{fig:sparch}) does not describe extremely long or short signal peptides. Theoretically, HMMs that describe the atypical signal peptides could be developed to consider also unusual structures, but such probabilistic frameworks have not still been implemented.

\begin{figure}[ht]\centering
\includegraphics[width=0.95\textwidth]{figures/reglen.png}
\caption{A) Distribution of lengths of signal peptides. B) Distribution of the lengths of signal peptide regions. The measure of the length is a number of amino acids in the subsequence. The data was extracted from 2~589 signal peptide sequences derived from UniProt database (see \textbf{Data selection} in \textbf{Methods}).}
\label{fig:reglen}
\end{figure}

Another feature, shared by majority of the signal peptide predicting software, is the orthogonal encoding of amino acids, in which a vector of 20 digits represents every amino acid. This method of encoding, however, does not preserve relationships between amino acids. The distance between residues is the same regardless of the similarity of their physicochemical properties. Such behaviour is especially undesired in signal peptides models because their regions are defined by specific features of amino acid residues and not by the simple occurrence of specific amino acids. In addition to this, such sparse encoding enforces larger data sets, which hinders their management and analysis~\citep{2002linamino}.

All programs used in signal peptide recognition are trained on real protein sequences. Therefore, they succeed in the recognition of peptides similar to those in the learning set, but fail in the case of artificial signal peptides. Such peptides are designed to increase effectiveness of protein secretion~\citep{2010futatsumorisugaisignal}. They are especially important in industrial applications to increase yield of proteins. Therefore, only explicit knowledge about the organization of signal peptides allows creating sequences that will be the most efficient in the export of proteins~\citep{2013ngengineering}. Signal peptides have also an important application in gene therapy. Mimicking the natural mechanism of protein export, artificial signal peptides with tumor epitopes increase the antitumor immune response~\citep{2003heenhanced}. Such epitopes must be properly inserted into a signal peptide without decreasing its secretion properties through disruption of the regional structure. Instead of time-consuming and expensive laboratory experiments, it would be very useful to survey in silico many artificial peptides to select the ones that would fulfil the designed role.

\section*{Methods}

\subsection*{Overview}

The functionality of a signal peptide depends not on exact sequence of specific amino acids, but on the physicochemical properties of residues in a given region. Henceforth, the usage of raw amino acid sequences is superfluous and introduces unnecessary information. To utilize this property of signal peptide recognition, we clustered amino acids into several groups based on the physicochemical features of residues.

The pre-processed sequences were further analyzed by the heuristic algorithm, which determines borders between three characteristic signal peptide regions, the enhanced version of algorithm presented in~\cite{1998nielsenprediction}. We refined some region recognition criteria to attune the algorithm to less typical signal peptides.

Next, two models are trained to recognize proteins with and without a signal peptide. The first one is a hidden semi-Markov model, in which each of three signal peptide regions is represented by a different hidden state. The additional fourth hidden state represents mature protein. Each state is described by the frequencies of amino acid groups  within that state. The distribution of hidden states durations, the number of amino acids related to each hidden state in signal peptide, is based on the empirical density of region lengths from the training set. Furthermore, the hidden semi-Markov model was enriched with n-grams representing signal peptide cleavage sites.

The second model is a simple probabilistic approach in which no association between amino acids was assumed, and probability of amino acids groups occurrence was determined by their frequencies in mature proteins.

\subsection*{Data selection}

Eukaryotic protein sequences and their annotations were properly prepared according to the literature of the subject and downloaded from UniProt database release 2015\_06. The positive set contained 2~589 sequences with an experimentally confirmed signal peptide, its start and cleavage site. Sequences with more than one cleavage site were excluded from the final data set. The negative set comprised 152~272 sequences without any signal peptide annotation. Protein sequences with ambiguous symbols: X, J, Z and B were removed from the final sets. Proteins with selenocysteine (U) were also excluded from data set, because there are no records of signal peptides containing this amino acid.

\subsection*{Clustering of amino acids}

It is worth to note, that signalHsmm is not a first software to use amino acid encoding in signal peptide prediction. BLOMAP~\citep{maetschke2005blomap} also employed similar strategy, but considered only encodings based on the substitution matrices. Substitution matrices are computed using protein sequences and knowing how atypical signal peptides are, we had chosen different approach.

Instead amino acids were clustered using several criteria relevant for the architecture of signal peptide: hydrophobicity, frequency in alpha-helices, polarity and size. High hydrophobicity is a determinant of the h-region, the core of signal peptides. Alpha-helix, the secondary structure of the h-region, is probably induced by the positively charged n-region. High polarity as well as smaller size are important features of the residues in the cleavage site~\citep{1994palzkillselection}.

\begin{table}[ht]
\centering
\caption{Properties used in clusterization.} 
\begin{tabular}{ll}
  \toprule
Criterion name & Property name \\ 
  \midrule
Size & Size~\citep{dawson1972size} \\ 
   \rowcolor[gray]{0.85}Size & Molecular weight~\citep{fasman1976handbook}\\ 
  Size & Residue volume~\citep{1973goldsackcontribution} \\ 
   \rowcolor[gray]{0.85}Size & Bulkiness~\citep{1968zimmermanthe} \\ 
  Hydrophobicity & Normalized hydrophobicity scales for alpha-proteins~\citep{1992cidhydrophobicity} \\ 
   \rowcolor[gray]{0.85}Hydrophobicity & Consensus normalized hydrophobicity scale~\citep{1984eisenbergthreedimensional} \\ 
  Hydrophobicity & Hydropathy index~\citep{1982kytea} \\ 
   \rowcolor[gray]{0.85}Hydrophobicity & Surrounding hydrophobicity in alpha-helix~\citep{1980ponnuswamyhydrophobic} \\ 
  Polarity & Polarity~\citep{1974granthamamino} \\ 
   \rowcolor[gray]{0.85}Polarity & Mean polarity~\citep{1988radzickainfluences} \\ 
  Frequency in alpha-helices & Signal sequence helical potential~\citep{1982argosstructural} \\ 
   \rowcolor[gray]{0.85}Frequency in alpha-helices & Normalized frequency of N-terminal helix~\citep{chou1978prediction} \\ 
  Frequency in alpha-helices & Relative frequency in alpha-helix~\citep{1990prabhakaranthe} \\ 
   \bottomrule
\end{tabular}
\label{tab:aaprop}
\end{table}

We selected 13 properties from AAIndex database~\citep{2008kawashimaaaindex} (see Tab.~\ref{tab:aaprop}), each attributed to a single criterion. We established 96 permutations of the properties, where each permutation contains only one property associated with a given criterion. Considering all permutations of properties, we created 96 possible clusterings of amino acids using Euclidean distance and Ward's method. Further, we cut the clusterings to create four-group encodings. As expected, some encodings (31\%) were identical.

To compare encodings we performed a 5-fold cross-validation training a new instance of signalHsmm on every encoding. We created balanced data sets by subsampling a number of proteins without a signal peptide equal to the number of proteins with signal peptide.
The analysis was carried using also redundant encodings, but results are not shown (HSMM is a deterministic algorithm and predictions for identical groupings of amino acids were the same). The cross-validation was repeated 60 times. At first, we planned 160 repetitions to ensure with probability 0.95, that every protein without signal peptide was in the learning set, but very small variance of performance measures (for example see Tab.~\ref{tab:perfmeas}) allowed us to relax this condition wihout sacrificing credibility of cross-validation.

\subsection*{Hidden semi-Markov model}
Hidden semi-Markov model (HSMM) is an extension of hidden Markov model (HMM).
As such, it proved to yield better performance in many fields of bioinformatics including
gene recognition~\citep{Pachter02applicationsof}, protein secondary structure prediction~\citep{16571137}
or automatic detection of heartbeats~\citep{7043102}.
Let us first briefly describe the idea that lies behind HMM. 
Suppose you have a sequence of observations e.g. amino acids, and you are interested in understanding underlying cause of their occurrence. 
HMM aims to answer that question by assuming a specific, yet flexible, structure of the problem.
HMM consists of two stochastic processes. First is discrete Markov chain $X_{t=1}^T$ on the set of so called hidden states $\{S_1, \dots S_n\}$.
They are "the cause" of the observations. At every step $t$, hidden state might change according to transition matrix
$A= (a)_{i,j=1}^n$, where $a_{i,j} = \mathcal{P}(X_{t+1} = S_j | X_t = S_i)$. In our application hidden states are signal peptide regions.
Second process $E_{t=1}^T$ is observation process, defined on the set of possible observations $\{O_1, \dots, O_m\}$. 
Observations are assumed to occur independently. Their distribution depends only on hidden state which emits it. 
Probabilities are given by a matrix $B$, $b_{i,k} = \mathcal{P}(E_t = O_k | X_t = S_i)$.
In our case, observations are (degenerated) amino acids.
Main goal in signalHsmm is to find, for a given peptide, most probable regions boundaries. This is achieved with
Viterbi algorithm.
For a good reference on HMM see~\citep{1989rabinera}.

In regular HMM, hidden state duration -- number of observations emitted by the hidden state, has geometric distribution.
\citep{Durbin98biologicalsequence} shows how to extend it for some different distributions without significant increase in computational 
complexity. Similar ideas were used for signal peptide recognition, for example in~\citep{2004klla}. 
It is, however, still not flexible enough. Empirical regional length distributions (see Fig.~\ref{fig:reglen})
are difficult to capture in that way.

%rysunek pokazujący o co chodzi w HMM
\begin{figure}[h]
\centering
\tikzstyle{block} = [draw,shape=circle, top color=white, bottom color=green!50!red!70 ,minimum size=4em]
\tikzstyle{blockComp} = [draw,shape=circle, top color=white, bottom color=red!50!black!70 ,minimum size=2.5em]
% diameter of semicircle used to indicate that two lines are not connected
\def\radius{.7mm} 
\tikzstyle{branch}=[shape=rectangle, top color=white, bottom color=green!50!red!70 ,minimum size=3pt,inner sep=0pt]
\tikzstyle{branch2}=[shape=rectangle, top color=white, bottom color=red!50!black!70 ,minimum size=3pt,inner sep=0pt]
\def\n{11}
 \tikzstyle{line} = [draw=black, color=black!30!white!50, line width=1.5mm, -latex'] 
 \tikzstyle{line2} = [draw=black, color=red!30!black!50, line width=1.5mm, -latex']
\tikzstyle{frame}=[text=black,above, bottom color=white, top color=blue!50!black!70  ]
\tikzstyle{frameComp}=[text=black,below, bottom color=white, top color=red!50!black!70  ]
 \tikzstyle{line3} = [color=green!50!red!70, line width=1.5mm]
\def\names{{"$O_{1,1}$","...","$O_{1,d_1}$", "...","$O_{k, 1}$", "...", "$O_{k, d_k}$"}}%
\begin{tikzpicture}[>=latex']
%wąsate nawiasy
\draw [decorate, decoration={brace,amplitude=10pt},line3,xshift=0pt,yshift=0pt] (2.5,1) -> (10.5,1) node [branch,midway,yshift=0.8cm,color=black] 
	{\textbf{Hidden states}};
% \draw [->,decoration={brace,mirror,amplitude=10pt},line2,xshift=0pt,yshift=0pt] (11,-3) -- (3,-3) node [branch2,midway,yshift=-0.6cm,color=black] 
% 	{\textbf{Kierunek odczytu na nici Cricka ($3' \to 5'$)}}; 
   % Draw blocks, inputs and outputs

\node[block, text width=3em] at (1+2,0) (block1) {\footnotesize 1st hidden state};
\node[blockComp] at (-.5+2,-2)  (komp1) {\footnotesize \pgfmathparse{\names[0]}\pgfmathresult};
\node[blockComp] at (1+2,-2)  (komp2) {\footnotesize \pgfmathparse{\names[1]}\pgfmathresult};
\node[blockComp] at (2.5+2,-2)  (komp3) {\footnotesize \pgfmathparse{\names[2]}\pgfmathresult};
\draw[line,->]  (block1) --  (komp1);
\draw[line,->]  (block1) --  (komp2);
\draw[line,->]  (block1) --  (komp3);
\draw[line3,->] (block1.east) -- +(2,0) node [branch,midway,yshift=0.2cm,xshift=-0.2cm,color=black] {\footnotesize \textbf{transition}};

\node[block, bottom color=green!50!red!80,] at (4.5+2,0) (block2) {...};
\draw[line3,->] (block2.east) -- +(2,0) node [branch,midway,yshift=0.2cm,xshift=-0.2cm,color=black] {\footnotesize \textbf{transition}};
\node[blockComp] at (4.5+2,-2)  (komp4) {\footnotesize \pgfmathparse{\names[3]}\pgfmathresult};
\draw[line,->]  (block2) --  (komp4) ;

\node[block,text width=3em, bottom color=green!50!red!90] at (8+2,0) (block3) {\footnotesize $k^{th}$ hidden state};
\node[blockComp] at (6.5+2,-2)  (komp5) {\footnotesize \pgfmathparse{\names[4]}\pgfmathresult};
\node[blockComp] at (8+2,-2)  (komp6) {\footnotesize \pgfmathparse{\names[5]}\pgfmathresult};
\node[blockComp] at (9.5+2,-2)  (komp7) {\footnotesize \pgfmathparse{\names[6]}\pgfmathresult};
\draw[line,->]  (block3) --  (komp5);
\draw[line,->]  (block3) --  (komp6);
\draw[line,->]  (block3) --  (komp7);

% \draw[<->, color=red!30!black!50, line width=1.5mm] (komp1) |- +(5,-1.5) node[line2, yshift=-0.5cm] {\textbf{Observations}} -| (komp7);

\draw [color=red!30!black!50, line width=1.5mm, decorate,decoration={brace,amplitude=10pt,mirror,raise=10pt},yshift=-10pt]
(1,-1.8) -- (5,-1.8) node [branch2, midway, yshift=-1cm,color=black] {\textbf{Duration of length $d_1$}};

\draw [color=red!30!black!50, line width=1.5mm, decorate,decoration={brace,amplitude=10pt,mirror,raise=10pt},yshift=-10pt]
(8,-1.8) -- (12,-1.8) node [branch2, midway, yshift=-1cm,color=black] {\textbf{Duration of length $d_k$}};

(1,-2.8) -- (12,-2.8) node [branch2, midway, yshift=-1.5cm,color=black] {\textbf{Observations, $\sum_{i=1}^k d_i = T$}};

\end{tikzpicture}
\caption{General scheme of hidden semi-Markov model.}
\label{fig:hsmm}
\end{figure}

The model we use is Hidden semi-Markov Model(HSMM)~\citep{Yu2010215}. It extends HMM by allowing 
any given hidden state duration distribution. Diagram of HSMM is presented in Fig.~\ref{fig:hsmm}.
In addition to matrices $A$ and $B$, model is given by probabilities of duration length in hidden states.
$$\mathcal{P}(\text{duration in state} = d | \text{state is } S_i), \;\; i = 1, \dots, n, \;\; d = 1, \dots, D$$
where $D$ is maximum allowed duration.
As our datasets are of reasonable size, $D$ is small -- around 30, computational effort is not much 
higher than with the regular HMM.

The model we use has a very specific structure. Hidden states are signal peptide regions.
Almost all entries in transition matrix $A$ are zeros, because regions are sequential.
Possible transitions are depicted as arrows in Fig.~\ref{fig:signalHsmm}.
Probabilities of observations for hidden states and hidden states durations were estimated from training data.
The advantage of HSMM model is not only a better performance but also its straightforwardness.
Fig.~\ref{fig:signalHsmm} is easy to interpret for a researcher without any mathematical background.

%model structure
\begin{figure}
 \begin{minipage}{.5\linewidth} 
     \resizebox{6.5cm}{!}{%
    \begin{tikzpicture}[->,>=stealth',shorten >=2pt,auto,node distance=4.5cm, thick]
    \label{fig:model_hsmm}
      \tikzstyle{line} = [draw=black, color=blue!30!black!50, line width=4.5mm, -latex']
      \tikzstyle{main node} = [circle,fill=blue!20,draw, minimum size = 2.2cm, font=\itshape,
         align=center,  top color=white, bottom color=blue!50!black!70 ] %font=\sffamily\small\bfseries,
      %nodes
      \node[main node]            (start') 	[]						{Start};	     
      \node[main node, bottom color=purple!70!black!70] 	(nregion') 	[right of=start',xshift=-5mm, yshift=15mm] 	{n-region};
      \node[main node, bottom color=pink!70!black!70] 	(hregion') 	[right of=nregion',xshift=-5mm,yshift=15mm] 	{h-region};
      \node[main node, bottom color=gray!70!black!70] 	(cregion') 	[right of=hregion',xshift=-5mm,yshift=-15mm] 	{c-region};
      \node[main node, bottom color=green!70!black!70] 	(mature') 	[right of=cregion',xshift=-5mm, yshift=-15mm] 	{Mature protein};

      %lines
      \path [line] (start')   edge node [left, color=black] {} (nregion');
      \path [line] (nregion') edge node [below, color=black] { } (hregion');
      \path [line] (hregion') edge node [below, color=black] { } (cregion');
      \path [line] (cregion') edge node [left, color=black] { } (mature');
      \draw [line] (start') to[out=340,in=200] (mature');
    \end{tikzpicture} }
    \caption{The diagram of signalHsmm.}
    \label{fig:signalHsmm}
\end{minipage}
 \begin{minipage}{.5\linewidth}
     \resizebox{6.5cm}{!}{%
    \begin{tikzpicture}[->,>=stealth',shorten >=2pt,auto,node distance=4.5cm, thick]
    \label{fig:model_hsmm_kMer}
      \tikzstyle{line} = [draw=black, color=blue!30!black!50, line width=4.5mm, -latex']
      \tikzstyle{main node} = [circle,fill=blue!20,draw, minimum size = 2.2cm, font=\itshape,
         align=center,  top color=white, bottom color=blue!50!black!70 ] %font=\sffamily\small\bfseries,
      %nodes
      \node[main node]            (start')   []  					{Start};	     
      \node[main node, bottom color=purple!70!black!70] 	(nregion') 	[right of=start',xshift=-25mm, yshift=35mm] 	{n-region};
      \node[main node, bottom color=pink!70!black!70] 	(hregion') 	[right of=nregion',xshift=-25mm,yshift=35mm] 	{h-region};
      \node[main node, bottom color=gray!70!black!70] 	(cregion') 	[right of=hregion',xshift=-10mm,yshift=-25mm] 	{c-region};
      \node[main node, bottom color=green!70!black!70] 	(mature') 	[below of=hregion',xshift=15mm, yshift=-90mm] 	{Mature protein};

      \node[main node, bottom color=yellow!70!black!70]   (kmer1) 	[below of=cregion',xshift=35mm,yshift=30mm] 	{Aminoacid 1};
      \node[main node, bottom color=yellow!70!black!70]   (kmer2) 	[below of=kmer1,xshift=-5mm,yshift=5mm] 	{Aminoacid 2};
      \node[main node, bottom color=yellow!70!black!70]   (kmer3) 	[below of=kmer2,xshift=-5mm,yshift=5mm] 	{Aminoacid 3};
      %lines
      \path [line] (start')   edge node [left, color=black] {} (nregion');
      \path [line] (nregion') edge node [below, color=black] { } (hregion');
      \path [line] (hregion') edge node [below, color=black] { } (cregion');
      \path [line] (cregion') to[out=230,in=90] (mature');
      \draw [line] (start') to[out=300,in=180] (mature');

      \draw [line] (cregion') edge node [below, color=black] { } (kmer1);
      \draw [line] (kmer1) edge node [left, color=black] {} (kmer2);
      \draw [line] (kmer2) edge node [left, color=black] {} (kmer3);
      \draw [line] (kmer3) edge node [left, color=black] {} (mature');

      %\draw [decorate,line width=3pt,decoration={brace,amplitude=10pt}] ([xshift=5mm, yshift=5mm]kmer1) -- ([yshift=-25mm]kmer3) node [black,midway,xshift=0.6cm] {\footnotesize $P_1$};
      \draw[decoration={brace,amplitude=10pt, raise=5pt},decorate, line width=2pt]  
      ([yshift=1mm]kmer1.east) -- ([yshift=-5mm]kmer3.east) node [black,midway,xshift=0.6cm] {\large \color{black!90} n-gram};
    \end{tikzpicture} }
    \caption{The diagram of signalHsmm extended with the n-gram cleavage site model.}
    \label{fig:ngramext}
    \end{minipage}
\end{figure}

\subsection*{n-gram extension}

Hidden semi-Markov models may be flexibly enhanced by adding additional hidden states. To improve our model, we added few supplementary states representing specific motives that might occur in proximity of cleavage site. The structure of cleavage sites, more conservative  than other parts of signal peptide~\citep{2004hillerpredisi}, may be mirrored by n-grams (k-mers), short vectors of $n$ characters derived from input sequences.

Using the biogram software~\citep{biogramPackage} we extracted n-grams from cleavage sites of signal peptides. The analyzed sequences were already encoded using amino acid classification providing the best sensitivity of a general model (see Tab.\ref{tab:best}. Selected n-grams, representing less common cleavage site motifs, were included in the HSMM model as the alternative paths at the end of c-region. The architecture of extended HSMM can be seen in Fig.~\ref{fig:ngramext}.

\section*{Results and discussion}

\subsection*{Cross-validation}

\begin{figure}[ht]\centering
\includegraphics[width=0.95\textwidth]{figures/cvres.png}
\caption{Sensitivity and specificity of encodings after cross-validation. 1. The encoding providing the best sensitivity (AUC = 0.9683, MCC = 0.8677). 2. The encoding providing the best specificity (AUC = 0.9338, MCC = 0.6474).}
\label{fig:cvres}
\end{figure}

\begin{table}[ht]
\begin{minipage}{.5\linewidth} 
\centering
\caption{The best sensitivity (final) encoding.} 
\begin{tabular}{l}
  \toprule
Groups \\ 
  \midrule
D, E, H, K, N, Q, R \\ 
   \rowcolor[gray]{0.85}G, P, S, T, Y \\ 
  F, I, L, M, V, W \\ 
   \rowcolor[gray]{0.85}A, C \\ 
   \bottomrule
\end{tabular}
\label{tab:best}
\end{minipage}
\begin{minipage}{.5\linewidth} 
\centering
\caption{The worst sensitivity encoding.} 
\begin{tabular}{l}
  \toprule
Groups \\ 
  \midrule
A, E, K, Q, R \\ 
   \rowcolor[gray]{0.85}D, G, N, P, S, T \\ 
  C, H, I, L, M, V \\ 
   \rowcolor[gray]{0.85}F, W, Y \\ 
   \bottomrule
\end{tabular}
\label{tab:worst}
\end{minipage}
\end{table}

\begin{figure}[ht]\centering
\includegraphics[width=0.95\textwidth]{figures/enccomp.png}
\caption{A comparison of encodings. A) Properties of amino acids in the best and the worst sensitivity encoding. B) Frequencies of amino acids belonging to the encodings in regions of signal peptide and mature proteins.}
\label{fig:enccomp}
\end{figure}

We used four performance measures to evaluate results of cross-validation: specificity, sensitivity, Matthew's Correlation Coefficient ($\phi$ coefficient) and Area Under the Curve (AUC). All encodings provide very good AUC (0.93 -- 0.97) and specificity (0.92 -- 0.96). The classification of amino acids has the biggest impact on sensitivity, which ranges from 0.66 to 0.94. The final signalHsmm algorithm uses the encoding, that yields the highest specificity and Matthew's Correlation coefficient as well as the second best AUC (see Fig.~\ref{fig:cvres}).

\begin{table}[ht]
\centering
\caption{Performance measures for the best encoding. 60 repetitions of cross-validation.} 
\begin{tabular}{lrr}
  \toprule
Measure & Mean & SD \\ 
  \midrule
AUC & 0.9683 & 0.0024 \\ 
   \rowcolor[gray]{0.85}MCC & 0.8677 & 0.0049 \\ 
  Sensitivity & 0.9406 & 0.0008 \\ 
   \rowcolor[gray]{0.85}Specificity & 0.9269 & 0.0050 \\ 
   \bottomrule
\end{tabular}
\label{tab:perfmeas}
\end{table}

\subsection*{Comparison of encodings}

We examined more deeply encodings with the best and the worst sensitivity. In both cases Group 1 tends to contain polar amino acids typical with the average size of residuals. In the best encoding, all charged amino acids, both acids and bases (even as weak bases as histidine), also belong to this group. This group, nearly absent in h-region, provides very good distinction between parts of signal peptide for best encoding, because is nearly absent in the h-region. In the worst encoding, where polar and charged character of the Group 1 is not that explicit, the difference in frequency between regions is also less visible.

The amino acids belonging to the Group 2 have quite low probability of occurring in an alpha-helix, but are still very diverse, but all . Polar but uncharged serine and threonine neighbour with alyphatic glycine and proline. In the worst encoding the polar character of the group is emphasized by the addition of asparagine. The best grouping contains also highly hydrophobic tyrosine. Despite these differences, the frequency of Group 2 seems to be comparable between two encodings.

The both groupings have strongly non-polar and aliphatic group 3 (see Tab.~\ref{tab:best}, Tab.~\ref{tab:best} and Fig.~\ref{fig:enccomp}) containing isoleucine, leucine, methionine and valine. Hyrophobic properties of this group in the best encoding are even more pronounced by the presence of tryptophan and phenylalanine. On the contrary, to the Group 3 in the worst encoding belong also more polar histidine. Hence the hydrophobic character of this group, it is dominant over the h-region in case of both amino acid classifications.

The fourth group is the most diverse among both encodings. The key is obvious in a case of the worst encoding. Phenylalanine, tryptophan and tyrosine are aromatic, highly hydrophobic amino acids. In contrast, the Group 4 in the best encoding contains only alanine and cysteine, which both are rather small amino acids and tend to appear in alpha-helices. This very unique composition seems to be typical for the c-region of signal peptide.

The encoding of amino acid plays crucial role in the recognition of signal peptide, but does not affect identification of proteins without a signal peptides (compare change in specificity and sensitivity on Fig.~\ref{fig:cvres}). The mature protein state in the HSMM model tends to have more uniform distribution of residues than signal peptide region, which make it more resistant to changes in amino acid groupings.

\subsection*{Benchmark test}

To provide the honest comparison, we trained our model on 2311 signal peptide-containing sequences deposited in Uniprot until 2010 year (an iteration of signalHsmm called signalHsmm-2010). The data set should correspond to the set used to train SignalP 4.1. Interestingly, our algorithm performed very well even if it was trained on a very small sets including only 336 sequences collected till 1987 year, just after the first method predicting signal peptide was published~\citep{1986vonheijnea}. The signalHsmm-1987 was able to very accurately predict a signal peptide, even better than predictors trained on richer data sets. It indicates that signalHsmm is very stable and is able to recover the structure of signal peptides from even very small data sets~\ref{tab:bench2010}.

\begin{table}[ht]
\centering
\caption{Comparison of Area Under the Curve, Sensitivity, Specificity and Matthews Correlation Coefficient for different classifiers.} 
\begin{tabular}{lrrrr}
  \toprule
Software name & AUC & Sensitivity & Specificity & MCC \\ 
  \midrule
signalP 4.1 (no tm) \citep{2011petersensignalp} & 0.9424 & 0.9720 & 0.9128 & 0.8859 \\ 
   \rowcolor[gray]{0.85}signalP 4.1 (tm) \citep{2011petersensignalp} & 0.9675 & 0.9579 & 0.9771 & 0.9353 \\ 
  PrediSi \citep{2004hillerpredisi} & 0.8959 & 0.9065 & 0.8853 & 0.7919 \\ 
   \rowcolor[gray]{0.85}Phobius \citep{2004klla} & 0.9515 & 0.9673 & 0.9358 & 0.9033 \\ 
  Philius \citep{2008reynoldstransmembrane} & 0.9376 & 0.9533 & 0.9220 & 0.8755 \\ 
   \rowcolor[gray]{0.85}signalHsmm-2010 & 0.9429 & 0.9393 & 0.8761 & 0.8166 \\ 
  signalHsmm-1989 & 0.9455 & 0.9486 & 0.8945 & 0.8440 \\ 
   \bottomrule
\end{tabular}
\label{tab:bench2010}
\end{table}

To check universality of our probabilistic model, we also validated it on signal peptides belonging to specific taxonomic groups. As the example, we chose the Plasmodiidae family including malaria parasites, because of her medical significance. The testing data set extracted from UniProt data base contained 102 proteins with a signal peptide and 358 without it. We repeated the benchmark scheme for different classifiers (Tab.~\ref{tab:bench2010plas}).

\begin{table}[ht]
\centering
\caption{Comparison of Area Under the Curve, H-measure and Matthews Correlation Coefficient for different classifiers considering only proteins belonging to Plasmodiidae.} 
\begin{tabular}{lrrrr}
  \toprule
Software name & AUC & Sensitivity & Specificity & MCC \\ 
  \midrule
signalP 4.1 (no tm) \citep{2011petersensignalp} & 0.7928 & 0.6471 & 0.9385 & 0.6185 \\ 
   \rowcolor[gray]{0.85}PrediSi \citep{2004hillerpredisi} & 0.6597 & 0.3725 & 0.9469 & 0.4028 \\ 
  Phobius \citep{2004klla} & 0.7963 & 0.6765 & 0.9162 & 0.5991 \\ 
   \rowcolor[gray]{0.85}Philius \citep{2008reynoldstransmembrane} & 0.7753 & 0.6176 & 0.9330 & 0.5841 \\ 
  signalHsmm-2010 & 0.9340 & 0.7941 & 0.8939 & 0.6526 \\ 
   \rowcolor[gray]{0.85}signalHsmm-1989 & 0.9326 & 0.9216 & 0.8966 & 0.7531 \\ 
   \bottomrule
\end{tabular}
\label{tab:bench2010plas}
\end{table}

signalHsmm has much better performance that other algorithms on proteins belonging to organisms from Plasmodiidae family. Since our learns more general decision rules, it is able to grasp the similarities between very variable sequences. Without losing the ability to predict the more common cases, signalHsmm is able to also deal well with very atypical signal peptides.

\subsection*{Conclusions}

The architecture of existing signal peptide predicting software is usually opaque, preventing analysis of decision rules included in the model. Without ability to trace the computations, researchers cannot discover which parameters are responsible for predictions. Considering the biological context of the problem, we see a need for a transparent model of signal peptide. signalHsmm is a first step in this direction, providing users with both prediction results and the extandable, readable predictive model.

Aside from a confirmation of the accepted signal peptide model (hydrophibicity of the h-region, polarity of the n-region), we also discovered new relationships. We found the alanine is one of the most typical amino acids for the c-region and occurs there far more often than in the other parts of signal peptides and mature proteins. The presence of alanine in the signal peptide cleavage site~\cite{1986vonheijnea} cannot explain such over-representation of group containing alanine in the c-region.

The good performance of signalHsmm confirms that properties of signal peptides do not depend on their exact sequence, but on the physicochemical features of the amino acids. Although this finding may seem to offer a minimal improvement when we consider the most popular signal peptides, this generalization pays off when we analyze more specific cases, as signal peptides of \textit{Plasmodiidae}.

The flexibility and efficient information recovery makes our model unique among similar software. The signalHsmm not only shows very good performance but is also adjustable enough to model properly very specific signal peptides belonging to poorly represented taxonomic groups in databases. Morrover, our method can effectively extract information from very small data sets, which in future may lead to new predictors, able to recognize less known signaling sequences.

\subsection*{Availability and implementation}
The signalHsmm prediction web-server is avaible under address: \url{http://smorfland.uni.wroc.pl/signalhsmm}.

signalHsmm is implemented as an R package freely available on CRAN: \url{http://cran.r-project.org/web/packages/signalHsmm}. Stand-alone version, in addition to the prediction of signal peptides, offers tools to build, train and test novel signal peptide models.

\section*{Acknowledgments}
PS and PB devised the HSMM model of signal peptide. MB and PS co-wrote the software. MB gathered data and carried the analysis.
MB, PS and PM drafted the manuscript. All authors revised the manuscript and approved the final version.

\bibliography{lokalizom}

\end{document}