### 2015-07-06 signalHSMM_kMer_report.Rmd
now report is in concordance with the new signalHsmm

---
title: "signalHsmm with k-mer hidden states"
author: "Piotr Sobczyk"
date: "19 czerwca 2015"
output: html_document
---

We shall compare performance of currently used software with signalHSMM trained on data obtained by 1989, by 2010,
with and without k-mer hidden state.

```{r, message=FALSE}
library(dplyr)
library(biogram)
library(signalHsmm)
library(seqinr)
library(hmeasure)
library(ggplot2)
library(knitr)
library(foreach)
library(doMC)
```

First let us load data
```{r, cache=TRUE}
other_soft <- read.csv2("benchmark_other.csv")

group_best <- list(`1` = c("r", "n", "d", "q", "e", "h", "k"), 
     `2` = c("g", "p", "s", "t", "y"), `3` = c("i", "l", "m", "f", "w", "v"), 
     `4` = c("a", "c"))

signalHsmm2010 <- train_hsmm(read_uniprot("sp1950_2010.txt", euk = TRUE, what = "signal"), 
                             aa_group = group_best)
signalHsmm1989 <- train_hsmm(read_uniprot("sp1950_1989.txt", euk = TRUE, what = "signal"), 
                             aa_group = group_best)
signalHSMM_proteins <- read_uniprot("sp1950_2010.txt", euk = TRUE, what = "signal")
benchmark_full <- read_uniprot("sp2010_2015.txt", euk = TRUE, what = "signal")
benchmark_data <- read.fasta("benchmark_data.fasta", seqtype = "AA")
benchmark_data2 <- lapply(c(benchmark_full,benchmark_data[-c(1:214)]), function(i) i[1L:ifelse(length(i) > 80, 80, length(i))])

# usuniecie 4 obserwacji negatywnych
negs <- c(432, 215, 280, 417)
benchmark_data2 <- benchmark_data2[-negs]
other_soft <- other_soft[-negs,]
real_labels <- c(rep(1, 214), rep(0, 214))
```

Now let us compare predictions on that data
```{r, warning=FALSE, cache=TRUE}
all_preds <- data.frame(other_soft,
                       signalHsmm2010 = pred2df(predict(signalHsmm2010, benchmark_data2))[["sp.probability"]],
                       signalHsmm1989 = pred2df(predict(signalHsmm1989, benchmark_data2))[["sp.probability"]])

bench_metrics <- HMeasure(real_labels, all_preds,
                          threshold = c(rep(0.5, 5), 0.1, 0.1))[["metrics"]]
kable(bench_metrics)
```

Ok, let's move to defining k-mer states.

Show me what you got
```{r}
b <- lapply(signalHSMM_proteins, function(x){
  cut <- attr(x, "sig")[2]
  deg_sample <- as.numeric(degenerate(toupper(x)[1L:min(70,length(x))], group_best))
  #remove atypical amino acids
  deg_sample <- na.omit(deg_sample)
  deg_sample[(cut-2):(cut+2)]
})

b <- matrix(unlist(b), ncol=5, byrow = TRUE)
ggplot(data.frame(table(apply(b, 1, function(x) paste0(x[2:4], collapse="")))),
       aes(x=Var1, y=Freq)) + geom_point()
ggplot(data.frame(table(apply(b, 1, function(x) paste0(x[1:3], collapse="")))),
       aes(x=Var1, y=Freq)) + geom_point()
```

From that we will extract some most promising k-mers

```{r}
p4 <- 0.1
p3 <- 0.25/(1-p4)
p2 <- 0.35/(1-p4)/(1-p3)
p1 <- 0.1/(1-p4)/(1-p3)/(1-p2)

sum(b[,1]%in%c(3) & b[,2]%in%c(1) & b[,3] %in% c(2) & b[,4]%in%c(1) & b[,4]%in%c(1,3))/nrow(b)
kMer1 <- list(3, 1, 2, 1, c(1,3))
pState1 <- 3
nState1 <- 4
pTrans1 <- p1

# -3|222
sum(b[,1]%in%c(2,3) & b[,2]%in%c(1,2,3) & b[,3] %in% c(2,4))/nrow(b)
kMer2 <- list(c(2,3),c(1,2), c(2,4))
pState2 <- 3
nState2 <- 4
pTrans2 <- p2

#-3|4(1,2,3,4)(2,4)
sum(b[,1]%in%c(4) & b[,2]%in%c(1,2,3) & b[,3] %in% c(2,4))/nrow(b)
kMer3 <- list(4, c(1,2,3), c(2,4))
pState3 <- 3
nState3 <- 4
pTrans3 <- p3

# kMer4
sum(b[,1]%in%c(2) & b[,2]%in%c(1) & b[,3] %in% c(1,4) & b[,4]%in%c(1,2,3,4))/nrow(b)
kMer4 <- list(2, 1, c(1,4))
pState4 <- 3
nState4 <- 4
pTrans4 <- p4

lastState1 = 4+1+length(kMer1)
lastState2 = lastState1+1+length(kMer2)
lastState3 = lastState2+1+length(kMer3)
lastState4 = lastState3+1+length(kMer4)
```

Now let us build a new model on **signalHsmm2010**
```{r}
  parametersSet2 <- add_k_mer_state(kMer1, signalHsmm2010$pipar, signalHsmm2010$tpmpar, signalHsmm2010$od, 
                                     signalHsmm2010$params, pState1, nState1, pTrans1, d=3)
  parametersSet3 <- add_k_mer_state(kMer2, pipar = parametersSet2$pipar, tpmpar = parametersSet2$tpmpar, 
                                     od = parametersSet2$od, params = parametersSet2$params,
                                     pState2, nState2, pTrans2, d=3)
  parametersSet4 <- add_k_mer_state(kMer3, pipar = parametersSet3$pipar, tpmpar = parametersSet3$tpmpar, 
                                     od = parametersSet3$od, params = parametersSet3$params, 
                                     pState3, nState3, pTrans3, d=3)
  parametersSet5 <- add_k_mer_state(kMer4, pipar = parametersSet4$pipar, tpmpar = parametersSet4$tpmpar, 
                                     od = parametersSet4$od, params = parametersSet4$params, 
                                     pState4, nState4, pTrans4, d=3)
  overall_probs_log <- signalHsmm2010$overall_probs_log
  maxSignal <- 45
```

```{r}
registerDoMC(cores=7)
n <- length(benchmark_data2)
results <- foreach(i=1:n) %dopar% {
    prot <- benchmark_data2[[i]]
    deg_sample <- na.omit(as.numeric(degenerate(toupper(prot)[1L:min(maxSignal, length(prot))], group_best)))

    viterbi_res <- duration_viterbi(deg_sample-1, signalHsmm2010$pipar, signalHsmm2010$tpmpar, 
                                     signalHsmm2010$od, signalHsmm2010$params)
    viterbi_path <- viterbi_res[["path"]]+1
    c_site <- ifelse(any(viterbi_path == 4), 
                     max(which(viterbi_path == 3)), 
                     length(deg_sample))
    prob.signal <- viterbi_res[["viterbi"]][c_site, viterbi_path[c_site]]
    prob.non <- Reduce(function(x, y) x + overall_probs_log[y], deg_sample[1L:c_site], 0)
    prob.total <- exp(prob.signal - prob.non)

    viterbi_res5 <- duration_viterbi(deg_sample-1, parametersSet5$pipar, parametersSet5$tpmpar, 
                                      parametersSet5$od, parametersSet5$params)
    viterbi_path5 <- viterbi_res5[["path"]]+1
    c_site5 <- ifelse(any(viterbi_path5 == 4), 
                      min(which(viterbi_path5 == 4))-1, 
                      length(deg_sample))
    c_site5 <- ifelse(any(viterbi_path5 == lastState1), 
                      min(which(viterbi_path5 == lastState1)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState2), 
                      min(which(viterbi_path5 == lastState2)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState3), 
                      min(which(viterbi_path5 == lastState3)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState4), 
                      min(which(viterbi_path5 == lastState4)), 
                      c_site5)
    #get probabilities of signal peptide model
    prob.signal5 <- viterbi_res5[["viterbi"]][c_site5, viterbi_path5[c_site5]]
    #get probabilities of no signal peptide model
    prob.non5 <- Reduce(function(x, y) x + overall_probs_log[y], deg_sample[1L:c_site5], 0)
    prob.total5 <- exp(prob.signal5 - prob.non5)

    c(1 - 1/(1 + prob.total), 1 - 1/(1 + prob.total5), c_site, c_site5)
}
probs <- t(sapply(results, function (x) head(x, 2)))
cut <- t(sapply(results, function (x) tail(x, 2)))
```

```{r, echo=FALSE}
ggplot(data.frame(noMer=probs[,1],kMer=probs[,2], sig = factor(real_labels)), aes(x=noMer,y=kMer, color=sig)) +
  geom_point() + xlab("No k-mer") + ylab("with k-mer") + 
  scale_color_discrete("Signal peptide") +geom_abline(slope=1)
```

```{r, echo=FALSE}
all_preds2 = all_preds
all_preds2$signalKmer = probs[,2]

bench_metrics2 <- HMeasure(real_labels, all_preds2,
                             threshold = c(rep(0.5, 5), 0.1, 0.1, 0.5))[["metrics"]]
kable(bench_metrics2)
```

```{r}
#zle <- which(real_labels==1 & probs[,2]<0.75)
zle <- which(real_labels==1 & probs[,2]<probs[,1] & probs[,2]<0.99)
b2 <- lapply(benchmark_full[zle], function(x){
  cut <- attr(x, "sig")[2]
  deg_sample <- as.numeric(degenerate(toupper(x)[1L:min(50, length(x))], group_best))
  deg_sample <- na.omit(deg_sample)
  deg_sample[(cut-2):(cut+2)]
})

b2 <- matrix(unlist(b2), ncol=5, byrow = TRUE)
kable(data.frame(probs[zle,], seq=apply(b2, 1, function(x) paste0(x, collapse="")), 
      csite=sapply(benchmark_full[zle], function(x) cut <- attr(x, "sig")[2])))

kable(data.frame(which(real_labels==0 & probs[,2]>0.5),
      probs[which(real_labels==0 & probs[,2]>0.5),1],
      probs[which(real_labels==0 & probs[,2]>0.5),2]))
```

Now, what happens when we use kernel density estimation?
```{r}
params = structure(c(0.00345173057742305216, 0.16303340280710845733, 0.18328936325298567089, 0.32191072063759601507, 0.12710730373689624328, 0.19479707117476749434, 0.07149501410718961392, 0.05102665325739586671, 0.02061953603483871952, 0.01232173478541280856, 0.00339291179532653139, 0.00314522392528060902, 0.00050543218629833602, 0.00000944936698432368, 0.00000000009143740374, 0.00000000000000000000, 0.00000000000000000004, 0.00000000000000000321, 0.00000000000000000290, 0.00000000000000000501, 0.00000000000000000835, 0.00000000000000000088, 0.00000000000000000146, 0.00000000000000000000, 0.00000000000000000592, 0.00000000000000000000, 0.00000000000000000345, 0.00000000000000000036, 0.00000000000000000132, 0.00000000000000000276, 0.00000000000000000001, 0.00000000000000000462, 0.00119575265523844728, 0.00135312474054827118, 0.00278142092346731536, 0.00184134527771419310, 0.00400555470679759759, 0.01737188174042058122, 0.04070639179659169687, 0.06492241958211328323, 0.10486627831232180064, 0.13185802303839422311, 0.16522286797419358240, 0.14092085504016543807, 0.13473723690401800956, 0.08898241528270439682, 0.05443355717845509839, 0.04403282913502165608, 0.02561338612799781561, 0.01320007003819387639, 0.01069067991482806901, 0.00517725596697340799, 0.00181848832708673882, 0.00093206572991894022, 0.00156817541687766086, 0.00051169053398325657, 0.00077312532731958154, 0.00004559742198773287, 0.00000001037808743568, 0.00000000000000871706, 0.00000000000000000000, 0.00000000000000000000, 0.00000000000000000000, 0.00000000000000000000, 0.00345173057742305216, 0.16303340280710845733, 0.18328936325298567089, 0.32191072063759601507, 0.12710730373689624328, 0.19479707117476749434, 0.07149501410718961392, 0.05102665325739586671, 0.02061953603483871952, 0.01232173478541280856, 0.00339291179532653139, 0.00314522392528060902, 0.00050543218629833602, 0.00000944936698432368, 0.00000000009143740374, 0.00000000000000000000, 0.00000000000000000004, 0.00000000000000000321, 0.00000000000000000290, 0.00000000000000000501, 0.00000000000000000835, 0.00000000000000000088, 0.00000000000000000146, 0.00000000000000000000, 0.00000000000000000592, 0.00000000000000000000, 0.00000000000000000345, 0.00000000000000000036, 0.00000000000000000132, 0.00000000000000000276, 0.00000000000000000001, 0.00000000000000000462, 0.03125, 0.03125, 0.03125,
0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125,
0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125,
0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125,
0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125, 0.03125,
0.03125), .Dim = c(32L, 4L),
.Dimnames = list(NULL, c("n", "h","c", "")))

  parametersSet2 <- add_k_mer_state(kMer1, signalHsmm2010$pipar, signalHsmm2010$tpmpar, signalHsmm2010$od, 
                                     params, pState1, nState1, pTrans1, d=3)
  parametersSet3 <- add_k_mer_state(kMer2, pipar = parametersSet2$pipar, tpmpar = parametersSet2$tpmpar, 
                                     od = parametersSet2$od, params = parametersSet2$params,
                                     pState2, nState2, pTrans2, d=3)
  parametersSet4 <- add_k_mer_state(kMer3, pipar = parametersSet3$pipar, tpmpar = parametersSet3$tpmpar, 
                                     od = parametersSet3$od, params = parametersSet3$params, 
                                     pState3, nState3, pTrans3, d=3)
  parametersSet5 <- add_k_mer_state(kMer4, pipar = parametersSet4$pipar, tpmpar = parametersSet4$tpmpar, 
                                     od = parametersSet4$od, params = parametersSet4$params, 
                                     pState4, nState4, pTrans4, d=3)
  overall_probs_log <- signalHsmm2010$overall_probs_log
  maxSignal <- 45
```

```{r}
registerDoMC(cores=7)
n <- length(benchmark_data2)
results <- foreach(i=1:n) %dopar% {
    prot <- benchmark_data2[[i]]
    deg_sample <- na.omit(as.numeric(degenerate(toupper(prot)[1L:min(maxSignal, length(prot))], group_best)))

    viterbi_res <- duration_viterbi(deg_sample-1, signalHsmm2010$pipar, signalHsmm2010$tpmpar, 
                                     signalHsmm2010$od, signalHsmm2010$params)
    viterbi_path <- viterbi_res[["path"]]+1
    c_site <- ifelse(any(viterbi_path == 4), 
                     max(which(viterbi_path == 3)), 
                     length(deg_sample))
    prob.signal <- viterbi_res[["viterbi"]][c_site, viterbi_path[c_site]]
    prob.non <- Reduce(function(x, y) x + overall_probs_log[y], deg_sample[1L:c_site], 0)
    prob.total <- exp(prob.signal - prob.non)

    viterbi_res5 <- duration_viterbi(deg_sample-1, parametersSet5$pipar, parametersSet5$tpmpar, 
                                      parametersSet5$od, parametersSet5$params)
    viterbi_path5 <- viterbi_res5[["path"]]+1
    c_site5 <- ifelse(any(viterbi_path5 == 4), 
                      min(which(viterbi_path5 == 4))-1, 
                      length(deg_sample))
    c_site5 <- ifelse(any(viterbi_path5 == lastState1), 
                      min(which(viterbi_path5 == lastState1)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState2), 
                      min(which(viterbi_path5 == lastState2)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState3), 
                      min(which(viterbi_path5 == lastState3)), 
                      c_site5)
    c_site5 <- ifelse(any(viterbi_path5 == lastState4), 
                      min(which(viterbi_path5 == lastState4)), 
                      c_site5)
    #get probabilities of signal peptide model
    prob.signal5 <- viterbi_res5[["viterbi"]][c_site5, viterbi_path5[c_site5]]
    #get probabilities of no signal peptide model
    prob.non5 <- Reduce(function(x, y) x + overall_probs_log[y], deg_sample[1L:c_site5], 0)
    prob.total5 <- exp(prob.signal5 - prob.non5)

    c(1 - 1/(1 + prob.total), 1 - 1/(1 + prob.total5), c_site, c_site5)
}
probs <- t(sapply(results, function (x) head(x, 2)))
cut <- t(sapply(results, function (x) tail(x, 2)))
```

```{r, echo=FALSE}
ggplot(data.frame(noMer=probs[,1],kMer=probs[,2], sig = factor(real_labels)), aes(x=noMer,y=kMer, color=sig)) +
  geom_point() + xlab("No k-mer") + ylab("with k-mer") + 
  scale_color_discrete("Signal peptide") +geom_abline(slope=1)
```

```{r, echo=FALSE}
all_preds2 = all_preds
all_preds2$signalKmer = probs[,2]

save(benchmark_data2, file = "clean_data.RData")

bench_metrics2 <- HMeasure(real_labels, all_preds2,
                             threshold = c(rep(0.5, 5), 0.1, 0.1, 0.5))[["metrics"]]
kable(bench_metrics2)
```