### 2015-08-03 literature/literature.Rmd
started reading literature

---
title: "Literature"
bibliography: "kmer_filo.bib"
output:
  html_document:
    toc: true
    theme: united
---

# One by one

```{r, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
library(knitcitations)
library(RefManageR)
cleanbib()
BibOptions(sorting = "none", check.entries = FALSE)
refs <- read.bibtex("kmer_filo.bib", .Encoding = "UTF-8")
```

```{r, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
cat(paste0("##", citet(refs[["saha_brief_2013"]])))
```

They mention few investigations, most of them uses n-grams. The `r citet(refs[["wang_application_2000"]])` was the most appealing.

```{r, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
cat(paste0("##", citet(refs[["wang_application_2000"]])))
```

Uses aggregating of amino acids (called exchange group). Do not state why they created such clusterings. 'Exchange groups represent conservative re- placements through evolution.' They also sport a 'distance measure', which is more like the IG in the QuiPT.

Cool idea: 'To compensate for the possible loss of information due to
ignoring the other 2-gram patterns, a linear correlation coefficient (LCC) between the values of the 436 2-gram patterns
with respect to the protein sequence S and the mean value
of the 436 2-gram patterns in the positive training dataset
is calculated and used as another input feature value for S.'

They have quite expanded motif usage for the bayesian neural network.

# References

```{r, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
sort(refs, sorting = "ynt")
```